Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-08-27 draft.tex
Revised Proposed Features figure and its label

Another observation during the analysis are very large decision trees achieving low error scores.
%%
Their complexity impose overfitting to the data and they are not suitable for extracting universal rules.
%%
\subsection{Proposed Features}
\label{sec:ProposedFeatures}
\begin{figure}[!h]
  %\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/placeholder}
  \caption{
	Proposed features.
	%%
	(a) dented vertebraes are a sign of heavy stress and bone absorption.
	%%
	(b) Spine canal thickness is associated with herniated disks.
	%%
	(c) the spine canal shape can be used to characterize scoliosis and lordosis as well as the \emph{Cobb-Angles} \cite{Cobb}.
	%%
	}
  \label{fig:proposedFeatures}
\end{figure}
%%
\noindent Pathological deformation of the lumbar spine is usually the last stage and is associated with very strong lower back pain \com{Zitat Frau Hegenscheid fragen}.
%%
We have to capture earlier signs of pathological change in order to derive better predictive models.
%%

A very early sign of pathological deformation is the \emph{bone resorption in the center of a vertebra}.
%%
They change from being block-shaped to be dented in the center (Fig~\ref{fig:proposedFeatures}~a).
%%
This information can be obtained by segmenting the whole vertebra or the top and bottom point if each vertebra center.
%%
Another valuable variable would be the \emph{spine canal thickness} (Fig~\ref{fig:proposedFeatures}~b).
%%
Low spine canal thickness can be an indicator for an impending herniated disk.
%%
Both surface texture of the vertebra and thickness of the spine canal are used to diagnose herniated disks.
%%
The \emph{overall spine canal shape} is also of interested, since scoliosis and lordosis can be characterized more precise by deriving the Cobb angles from this shape (Fig~\ref{fig:proposedFeatures}~c).
%% ------------------------------------
\subsection{Applicability of the Method}
%%
Classification methods based on decision trees showed to be useful to assess the discriminative power of a variable set.
%%
Their ability to assess variable combination makes them more powerful than correlation coefficients calculated for each variable.
%%
This advantage comes with a much more complex output, making their results harder to assess and to abstract.
%%
Our method to plot derived metrics and custom tailored error measures proved to be effective.
%%
Huge result spaces could be navigated fast using our brushable visualization.
%%
Therefore the method is applicable not only for deriving information based on image data, but on all potential target variables.
%%

>>>>

Another observation during the analysis are very large decision trees achieving low error scores.
%%
Their complexity impose overfitting to the data and they are not suitable for extracting universal rules.
%%
\subsection{Proposed Features}
\label{sec:ProposedFeatures}
\begin{figure}[!h]
  %\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/proposed-features}
  \caption{
	Proposed features.
	%%
	(a) dented vertebraes are a sign of heavy stress and bone absorption.
	%%
	(b, c) the spine canal shape can be used to characterize scoliosis (curvature sagittal) and lordosis (curvature coronal) as well as the \emph{Cobb-Angles} \cite{Cobb}.
	%%
	(c) Spine canal thickness is associated with herniated disks.
	%%
	}
  \label{fig:proposedFeatures}
\end{figure}
%%
\noindent Pathological deformation of the lumbar spine is usually the last stage and is associated with very strong lower back pain \com{Zitat Frau Hegenscheid fragen}.
%%
We have to capture earlier signs of pathological change in order to derive better predictive models.
%%

A very early sign of pathological deformation is the \emph{bone resorption in the center of a vertebra}.
%%
They change from being block-shaped to be dented in the center (Fig~\ref{fig:proposedFeatures}~a).
%%
This information can be obtained by segmenting the whole vertebra or the top and bottom point if each vertebra center.
%%
Another valuable variable would be the \emph{spine canal thickness} (Fig~\ref{fig:proposedFeatures}~c).
%%
Low spine canal thickness can be an indicator for an impending herniated disk.
%%
Both surface texture of the vertebra and thickness of the spine canal are used to diagnose herniated disks.
%%
The \emph{overall spine canal shape} is also of interested, since scoliosis and lordosis can be characterized more precise by deriving the Cobb angles from this shape (Fig~\ref{fig:proposedFeatures}~b,~c).
%% ------------------------------------
\subsection{Applicability of the Method}
%%
Classification methods based on decision trees showed to be useful to assess the discriminative power of a variable set.
%%
Their ability to assess variable combination makes them more powerful than correlation coefficients calculated for each variable.
%%
This advantage comes with a much more complex output, making their results harder to assess and to abstract.
%%
Our method to plot derived metrics and custom tailored error measures proved to be effective.
%%
Huge result spaces could be navigated fast using our brushable visualization.
%%
Therefore the method is applicable not only for deriving information based on image data, but on all potential target variables.
%%
